---
title: "ridgereg"
author: "Boxi Zhang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(mlbench)
library(caret)
require(ridge)
library(methods)
data("BostonHousing")
```

```{r, echo=FALSE, eval=TRUE}
ridgereg <- setRefClass("ridgereg", 
                        fields = list(formula="formula",
                                      data="data.frame",
                                      reg_coe="matrix",
                                      fit_val="matrix",
                                      data_name="character",
                                      lambda="numeric"),
                        
                        methods = list(
                          initialize = function(formula, data, lambda=0){
                            
                            formula <<- formula
                            data <<- data
                            lambda <<- lambda
                            
                            data_name <<- deparse(substitute(data))
                            
                            #getting y variable from the formula
                            get_y <- all.vars(formula)[1]
                            
                            #defining X and y
                            y <- data[[get_y]]
                            X <- model.matrix(formula, data)
                            
                            #Normalize
                            p <- ncol(X)
                            X <- cbind(X[,1], scale(X[,-1]))
                            
                            #Ridge regression
                            # reg_coe <<- solve(t(X)%*%X+lambda*diag(p)) %*% t(X)%*%y
                            # dimnames(reg_coe) <<- list(c("(Intercept)", all.vars(formula)[-1]), NULL)

                            # fit_val <<- X%*%reg_coe
                            
                            #QR Decomposition
                            y<-as.matrix(data[get_y],ncol=1)
                            X_qr<-rbind(X,sqrt(lambda)*diag(p))
                            y_qr<-rbind(y,matrix(data = 0, nrow = p, ncol = 1))
                            
                            QR <- qr((X_qr))
                            Q <- qr.Q(QR)
                            R <- qr.R(QR)
                            
                            reg_coe <<- qr.solve(R) %*% t(Q) %*% as.matrix(y_qr)
                            dimnames(reg_coe) <<- list(c("(Intercept)", all.vars(formula)[-1]), NULL)
                            
                            fit_val <<- X %*% reg_coe
                          },
                          
                          results = function(){
                            results <- list()
                            
                            results$coef <- reg_coe
                            results$fitted <- fit_val
                            return(results)
                            
                          },
                          
                          print = function(){
                            "Print out the coefficients and coefficient names"
                            
                            cat("Call: \n ")
                            cat(paste0("ridgereg(formula = ",format(formula),", data = ",data_name,", lambda = ",lambda,")\n\n"))
                            cat("Coefficients: \n")
                            cat(" ",row.names(reg_coe), "\n    ", sep = "  ")
                            cat(t(reg_coe[,1]), sep="    ")
                            
                          },
                          
                          
                          predict = function(newdata=NULL){ #if newdata is used, it should be a data frame
                            "Prints out the predicted values or if newdata is used, prints out predicted values for the new data set"
                            
                            if(is.null(newdata)){
                              result <- structure(c(results()[[2]]), names=(1:length(results()[[2]])))
                            } else{
                              X<-model.matrix(object=formula, data=newdata)
                              X[,2:ncol(X)] <- scale(X[,-1])
                              result <- (X %*% results()[[1]])[,1]
                            }
                            return(result)
                          },
                          
                          # predict = function(){
                          #   "Return the predicted values y_hat"
                          #   return(fit_val)
                          # },
                          
                          coef = function(){
                            "Return the coefficients as a named vector"
                            return(reg_coe)
                          }
                        )
)
```

# Introduction
In this vignette, it will be shown how to do a simple prediction problem using ridgereg() function in the package "mlr" with the data _BostonHousing_ found in the mlbench package. The dataset consists of 506 observations of 14 variables. The median value of house price in $1000s, denoted by MEDV, is the outcome or the target variable in our model. Below is a brief description of each feature and the outcome in our dataset:

1. CRIM \-\-\- per capita crime rate by town
2. ZN \-\-\- proportion of residential land zoned for lots over 25,000 sq.ft
3. INDUS \-\-\- proportion of non-retail business acres per town
4. CHAS \-\-\- Charles River dummy variable (1 if tract bounds river; else 0)
5. NOX \-\-\- nitric oxides concentration (parts per 10 million)
6. RM \-\-\- average number of rooms per dwelling
7. AGE \-\-\- proportion of owner-occupied units built prior to 1940
8. DIS \-\-\- weighted distances to five Boston employment centres
9. RAD \-\-\- index of accessibility to radial highways
10. TAX \-\-\- full-value property-tax rate per $10,000
11. PTRATIO \-\-\- pupil-teacher ratio by town
12. B \-\-\- 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town
13. LSTAT \-\-\- % lower status of the population
14. MEDV \-\-\- Median value of owner-occupied homes in $1000’s

First 6 rows of the dataset are shown below:
```{r}
head(BostonHousing)
```


###1. Divide the BostonHousing data into a test and training dataset
```{r}
set.seed(12345)
inTrain <- createDataPartition(BostonHousing$medv, p = .7, list = FALSE)
training <- BostonHousing[ inTrain,]
testing  <- BostonHousing[-inTrain,]

#For the 10-fold cross-validation
ctrl <- caret::trainControl(method = "repeatedcv",
  number = 10, # k=10
  repeats = 10) # repeat 10 times
```
For future analysis, we standardize all predictors. The factor variable chas contains only two classes, so no dummies are needed and we may convert it into a usual numeric variable.

```{r}
BostonHousing$chas <- as.numeric(BostonHousing$chas) - 1
BostonHousing <- scale(BostonHousing)
```

###2-1. Fit a linear regression model.
```{r}
set.seed(12345)
lm <- caret::train(medv ~ ., 
            data = training,
            method = "lm",
            trControl = ctrl)
lm
```

###2-2. Fit a linear regression model with forward selection.
```{r}
set.seed(12345)
lflmGrid <- expand.grid(nvmax=1:(ncol(training)-1))
lflm <- caret::train(medv ~ ., 
              data = training,
              method = "leapForward",
              trControl = ctrl,
              tuneGrid = lflmGrid)

lflm

summary(lflm)
```

###3. Evaluate the performance of this model on the training dataset.
Based on the RMSE results above, regression model with 11 factors(excluding indus, age) is slightly better than with all 13 factors model. Though the difference is very small, model with 10 factors is simpler. For a detailed evaluation, residual vs fitted plot should be drawn.

```{r, fig.height= 7, fig.width= 7 }
p_data <- as.data.frame(cbind(predict(lflm), scale(resid(lflm))))

plot1 <- ggplot(p_data, aes(x = p_data[,1], y = p_data[,2])) + geom_point() +
          geom_smooth(method = "loess", color = "red") + xlab("Fitted") +
          ylab("Residuals") + ggtitle("Resid vs Fitted with 11 factors") +
          theme(plot.title = element_text(hjust = 0.5))

plot(plot1)
```

Above plot implies a clustering tendency of data and further investigation is needed to select a better model to fit this data. 

###4~5. Fit a ridge regression model using ridgereg() function in mlr package and finding a best parameter lambda using 10-fold cross-validation on the training set.
```{r, warning=FALSE}
#To use rigdereg() function
rglm <- list(type=c("Classification", "Regression"),
            library="ridge",
            loop=NULL,
            prob=NULL)

#Parameter setting
rglm$label<-"Ridge regression by mlr package"

rglm$parameters <- data.frame(parameter = "lambda",
                  class = "numeric",
                  label = "lambda")

rglm$fit <- function(x, y, lambda, param, lev, last, classProbs, ...) { 
  dat <- as.data.frame(x)
  
  respvector <- NULL
  respname <- NULL
  respnum <- NULL

  for(i in 1:ncol(x)){
    if(identical(y,dat[,i])){
    respvector <- dat[,i]
    respname <- names(x)[i]
    respnum <- i
    }
  }

  formula <- paste(respname,"~", sep="")

  if(ncol(x) > 1){
    for(i in 1:ncol(x)-1){
            formula <- paste(formula, "+", names(dat)[i], sep="")
    }
  }

  formula <- as.formula(formula)
  
  ridgereg$new(formula = formula, data=dat,lambda= param$lambda)
 }
 
rglm$predict <- function(modelFit, newdata, preProc=NULL, submodels = NULL){
  if (!is.data.frame(newdata)) 
    newdata <- as.data.frame(newdata)
  modelFit$predict(newdata)
}

rglm$sort  <- function(x) x[order(-x$lambda),]


rglm$grid <- function(x, y, len=NULL, search="grid"){
  data.frame(lambda=seq(from=0, to=5, by=1))
}

set.seed(12345)
rglm_fit <- caret::train(x = training,
                  y = training$medv,
                  method = rglm,
                  trControl = ctrl)

rglm_fit
```
With the caret package and ridgereg() function, ridge regression model is fitted to the training data. For the different values of lambda, 10-fold cross-validation is used to find the best hyperparameter value. As seen above, final value used for the model is lambda = 3. 

###6. Evaluate the performance of all three models
```{r}
#Linear regression
lm$results[,1:4]

#Regression with forward selection
lflm$results[11,1:4]

#Ridge regression
rglm_fit$results[4,1:4]
```
As shown above, regression with forward selection model has the lowest RMSE and MAE and highest Rsquared value. It means that this model is the best model among those 3 models. However, all 3 models show more or less the same values. In this case, most simple model could be selected. 

